---
title: DATA 606 Data Project Proposal
author: Chester Poon
---

### Data Preparation

```{r setup, echo=TRUE, results='hide', warning=FALSE, message=FALSE}
library(ggplot2)
library(dplyr)

# load data - We'll pull the variables we want. We'll keep a few extra in case we want to change our research question.
pd_killing <- read.csv('police_killings.csv')
pd_killing <- select(pd_killing,name,
                     raceethnicity,
                     tract_ce,
                     armed,
                     pop,
                     share_white,
                     share_black,
                     share_hispanic,
                     h_income
                     )

#Convert share to numeric proportion
pd_killing$share_white <- 
  as.numeric(levels(pd_killing$share_white))[as.integer(pd_killing$share_white)]/100
pd_killing$share_black <- 
  as.numeric(levels(pd_killing$share_black))[as.integer(pd_killing$share_black)]/100
pd_killing$share_hispanic <- 
  as.numeric(levels(pd_killing$share_hispanic))[as.integer(pd_killing$share_hispanic)]/100
pd_killing$tract_ce <- as.factor(pd_killing$tract_ce)

#2010 census population for blacks and whites
n_black <- 40250635
n_white <- 241937061

#prepare data for overall US stats for white victims and black victims per 1,000,000
b <- c('Black',(32/n_black)*1000000)
w <- c('White',(51/n_white)*1000000)
bw <- data.frame(t(data.frame(b,w)))
colnames(bw) <- c('race','rate')

#Subset and prepare the data we want for our research question.
pd_ksub <- pd_killing %>%
  filter(armed == 'No',
         raceethnicity %in% c('White','Black')) %>%
  select(raceethnicity,
         pop,
         share_white,
         share_black
         ) %>%
  mutate(white = share_white*pop,
         black = share_black*pop)
```


### Research question 

**You should phrase your research question in a way that matches up with the scope of inference your dataset allows for.**

I'm interested in examining two questions.

Are white Americans more likely to be killed by police within majority white census tracts? Are unarmed black Americans more likely to be killed by police anywhere when compared to unarmed white Americans?

### Cases 

**What are the cases, and how many are there?**

Each case in the data set represents one police killing. There are 467 observations in 2015. More specifically as it pertains to our question of unarmed victims, there are 83 observations (51 white, 32 black).

### Data collection 

**Describe the method of data collection.**

According to Ben Casselman of fivethirtyeight.com, data was compiled from the Guardian via "media coverage, reader submissions and open-sourced efforts". In addition, fivethirtyeight.com included census tracts to the dataset that I'll be using for this project.

For general nationwide population statistics, data was taken from the 2010 census.

### Type of study 

**What type of study is this (observational/experiment)?**

This is an observational study.

### Data Source 

**If you collected the data, state self-collected. If not, provide a citation/link.**

Data is collected by The Guardian and is available online here: 
<https://www.theguardian.com/us-news/ng-interactive/2015/jun/01/the-counted-police-killings-us-database>

Census tracts were added by fivethirtyeight.com to the 2015 data set. This is the primary source I will use for this project and is available online here: <https://fivethirtyeight.com/features/where-police-have-killed-americans-in-2015/>

2010 United States Census data is available here: <https://factfinder.census.gov/faces/tableservices/jsf/pages/productview.xhtml?src=bkmk>

Casselman, B., Fischer-Baum, R. (2015, June 3) *Where Police Have Killed Americans In 2015* Retrieved from https://fivethirtyeight.com/features/where-police-have-killed-americans-in-2015/

### Dependent Variable

**What is the response variable? Is it quantitative or qualitative?**

The response variable is quantitative: number/rate of killings

### Independent Variable

**You should have two independent variables, one quantitative and one qualitative.**

Independent variables are:

- race/ethnicity
- census tract/census tract population
- share_white
- share_black

### Relevant summary statistics 

**Provide summary statistics for each the variables. Also include appropriate visualizations related to your research question (e.g. scatter plot, boxplots, etc). This step requires the use of R, hence a code chunk is provided below. Insert more code chunks as needed.**

#####Summary statistics

```{r}
summary(pd_ksub)
```

#####Rate (Number of victims by race / National population by race per 1,000,000) of police killing by race.

```{r}
ggplot(bw, aes(x = race,y = rate)) +
  geom_bar(stat = 'identity')
```

#####Boxplot Analysis of victims by race and census tract "share" of population.

Noteable in that it appears that whites are more likely to be killed by police in predominantly white census tracts as expected, but blacks appear to be killed by police more often in predominantly non-black census tracts. 

```{r}
boxplot(pd_ksub$share_white,pd_ksub$share_black)
```

